---
title: "2022 Fantasy Football Wide Receiver Analysis"
author: "Matt Schmidt @WheyGood"
date: "2023-03-25"
output: pdf_document
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Load the data source csv file and all the libraries for the project
df <- read.csv("~/R_Data/WR_analysis_data.csv")

library(ggplot2)
library(dplyr)
library(devtools)
library(smplot2)
library(knitr)
library(stargazer)
#library(kableExtra)
#library(tinytex)

#Change the Theme (background) of the Charts
theme_set(theme_minimal())
```

# Introduction

Fantasy football is a statistics focused game based on the real life production
of professional National Football League players.  Success is found when the
user can more accurately predict the seasonal statistical output of said 
players than their opponents.  This report aims to identify specific advanced 
receiver stats that may aid in the prediction of future fantasy point output.  

Pearson correlation will be used along with regression analysis to determine 
which statistic will be most important to predict future fantasy success. 
The distribution of points and targets will also be examined to view outlier 
values which, in this case, are very much desired. 

## Pearson Correlation Analysis of Advanced WR Stats with Fantasy Point Output
```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.show="hold", out.width="50%"}

#Create the Air Yards vs Fantasy Points Correlation Graph
p1 <- ggplot(data=df, mapping=aes(x=air_yards, y=points)) + 
      geom_point(shape= 19, size=3, color='#311D00', alpha=.75) + 
      sm_statCorr(color='#FF3C00', separate_by = '\n')+
      labs(title="Air Yards and Fanasty Points Correlation", 
           x="Air Yards", y="Fantasy Points")
#Create the ADOT vs Fantasy Points Correlation Graph
p4 <- ggplot(data=df, mapping=aes(x=air_per_rec, y=points)) + 
  geom_point(shape= 19, size=3, color='#311D00', alpha=.75) + 
  sm_statCorr(color='#FF3C00', separate_by = '\n')+
  labs(title="Average Depth of Target and Fanasty Points Correlation", 
       x="ADOT (yards)", y="Fantasy Points")
#Create the Targets vs Fantasy Points Correlation Graph
p2 <- ggplot(data=df, mapping=aes(x=targets, y=points)) + 
  geom_point(shape= 19, size=3, color='#311D00', alpha=.75) + 
  sm_statCorr(color='#FF3C00', separate_by = '\n')+
  labs(title="Targets and Fanasty Points Correlation", 
       x="Targets", y="Fantasy Points")
#Create the Percent of Team Targets vs Fantasy Points Correlation Graph
p5 <- ggplot(data=df, mapping=aes(x=percent_team_targets, y=points)) + 
  geom_point(shape= 19, size=3, color='#311D00', alpha=.75) + 
  sm_statCorr(color='#FF3C00', separate_by = '\n')+
  labs(title="Percentage of Team Targets and Fanasty Points Correlation", 
       x="Percentage of Team Targets", y="Fantasy Points")

p1
p4
p2
p5


```

```{r, echo=FALSE}
table1 <- data.frame(Statistic = c("Air Yards", "Average Depth of Target", 
                                   "Targets", "Percentage of Team Targets")
                      ,Correlation_Coefficient = c(0.95, 0.013, 0.97, 0.95)
                      ,P_value = c(0.001, 0.86, 0.001, 0.001))

#kable(table1)
kable(table1)
```


##### Table 1: Correlation coefficents and p values of corresponding fantasy statistics



## Regression Analysis of Advanced WR Stats
```{r, echo=FALSE, message=FALSE, results='asis'}
# Create Smaller subset of wrs Database
dataset <- df %>% select(points, air_yards, air_per_rec, targets, 
                          percent_team_targets, red_zone_targets)

dataset$percent_team_targets_mult <- dataset$percent_team_targets * 100

reg1 <- lm(points ~ air_yards, data=dataset)
reg2 <- lm(points ~ air_per_rec, data=dataset)
reg3 <- lm(points ~ targets, data=dataset)
reg4 <- lm(points ~ percent_team_targets_mult, data=dataset)

stargazer(reg1, reg2, reg3, reg4, 
          title="Single Regression Results per Variable",
          type="latex",
          header=FALSE)

```



## Distribution of Points and Targets

```{r, echo=FALSE,message=FALSE,warning=FALSE, fig.show="hold", out.width="100%"}

dist1 <- ggplot(data = df, mapping = aes(x = " ", y = points)) + 
          sm_boxplot(shape=19, color='#FF3C00') + 
          ggtitle('Distribution of Points for all Players')

dist2 <- ggplot(data = df, mapping = aes(x = " ", y = targets)) + 
          sm_boxplot(shape=19, color='#FF3C00') + 
          ggtitle('Distribution of Targets for all Players')


dist1
dist2

```
\newpage

```{r, echo=FALSE, message=FALSE, fig.align="center"}
rank1 <- df[order(df$targets, decreasing = TRUE), 
             c('player', 'team', 'targets')] %>% top_n(20)

rank2 <- df[order(df$air_yards, decreasing = TRUE), 
             c('player', 'team', 'air_yards')] %>% top_n(20)

rank3 <- df[order(df$percent_team_targets, decreasing = TRUE), 
             c('player', 'team', 'percent_team_targets')] %>% top_n(20)

kable(list(rank2, rank3), booktabs=TRUE, align="llc", linesep = "",)
kable(rank1, booktabs=TRUE, align="llc", linesep = "")

```

\newpage

# Conclusion

The correlation and regression analysis of the 2022 wide receiver data
demonstrates the importance of usage in the offense. The targets and air yards 
a player receives most indicate fantasy football success. These variables both 
have significant correlation coefficients and with a large R squared value shows
when these variables increase, the points will follow.  

The distribution of wide receiver points shows a nice cluster of around six
elite scoring output-generating players.  These players should be valued highly
and selected early in fantasy drafts, as the next grouping of players many 
very similar outputs. Either take an high powered receiver early in drafts our
wait a couple rounds to select one of many with close target numbers.

Future work will be done to determine if one can predict future target 
share for a player in a specific offense.  This will require more team specific
coaching data with a sprinkling of quarterback data.  The trends in passing 
versus running plays for specific coaches will be important to examine for the 
prediction. 



